Sampling requirements to detect ecosystem change

### 2019-11-21 R/temporal-sampling.Rmd
edits on temporal sampling

@@

```{r read zurich data, include=T}
zh <- read.delim(paste0(getpath4data(),"ZH17_21_clado.txt")); is.zh=T
zh <- read.delim(paste0(getpath4data(),"VAR10-10-clado.txt")); is.zh=F
```

>>>>

- In paleo studies, if we cannot sample continously, we would sometimes chose to sample at regular intervals (e.g., every 5 cm). Would an iterative approach be more effective in detecting major changes?
- Limitations: right now we are detecting changes on a single vector (in ecological dataset, it would be for example a PC axis). We talked about how a user could sample, e.g., 5 samples, and then decide where to add some samples, depending on when the change was detected. However, the PC ordination will be done depending on the community structure. Are we including a bias by doing that?

@@

```{r read zurich data, include=T}
zh <- read.delim(paste0(getpath4data(),"ZH17_21_clado.txt")); is.zh=T
zh2 <- read.delim(paste0(getpath4data(),"VAR10-10-clado.txt")); is.zh=F
zh3 <- cbind(zh, matrix(rep(0,nrow(zh)*(ncol(zh)-1)), nrow(zh)))
zh3 <- rbind(zh3, cbind(matrix(c(1700:1710), rep(0,10*(ncol(zh)-1))),unlist(as.vector(zh2[1:10,-1]))) )
zh<-zh3
```
